borrow-lambda

### 2025-04-19 paper/paper.tex
some code examples

\subsection{Syntax}
Our syntax of terms is as follows.
\begin{center}
    \begin{bnf}
        $t$ : \textsf{Term} ::=
        | $x$ | $\lambda x\ldotp t$ | $t_1\ t_2$ | $\&x$ | $\&\text{mut}\ x$ | $*x$
        | if $t_1$ then $t_2$ else $t_3$
        | let $x = t_1$ in $t_2$
        | $x \coloneq t$
        | $*x \coloneq t$
        | true
        | false
        | unit
        | 0
        | pred $t$
        | succ $t$
        | iszero $t$
        | natvec($t_1,t_2,\dots t_n$)
        | get($t_1,t_2$)
        | getmut($t_1,t_2$)
        | push($t_1,t_2$)
        | pop($t_1$) 
    \end{bnf}
\end{center}
In addition to standard natural numbers and booleanss and some simple operations for them, we add
mutable and immutable references, the dereference operator, assignment and
dereference-assignment, as well as a vector of nats for a more complex
(non-copyable) data structure.

We add terms representing operations that create a vector given a list of terms, 
get a reference (mutable or immutable) of an element of a vector given a reference to the vector and an index, 
a similar method for pushing an element to a vector, and a method for popping an element
(and returning the actual element as opposed to just a reference of it). Note that
in this implementation, all variables are mutable.

@@

Lastly, when relevant, our typing rules make use of the following named set,
denoting types that are copyable.
$$\text{Copy}=\{\text{Nat}, \text{Bool}, \text{Unit},\&^\kappa_{\text{shr}} T\}$$

@@

This pass also does all the necessary work for lifetimes; assigning lifetimes
to each of the variables and their respective borrows, checking the subtype
relation of lifetimes, and handling instantiation of functions polymorphic over
lifetimes.

\texttt{TODO: add code examples and rules}

>>>>

\subsection{Syntax}
Our syntax of terms is as follows.
\begin{center}
    \begin{bnf}
        $t$ : \textsf{Term} ::=
        | $x$ | $\lambda x\ldotp t$ | $t_1\ t_2$ | $\&x$ | $\&\text{mut}\ x$ | $*x$
        | if $t_1$ then $t_2$ else $t_3$
        | let $x = t_1$ in $t_2$
        | $x \coloneq t$
        | $*x \coloneq t$
        | true
        | false
        | unit
        | 0
        | pred $t$
        | succ $t$
        | iszero $t$
        | natvec($t_1,t_2,\dots t_n$)
        | get($t_1,t_2$)
        | getmut($t_1,t_2$)
        | push($t_1,t_2$)
        | pop($t_1$) 
        | $t\ \colon T$
    \end{bnf}
\end{center}
In addition to standard natural numbers and booleans and some simple operations for them, we add
mutable and immutable references, the dereference operator, assignment and
dereference-assignment, as well as a vector of nats for a more complex
(non-copyable) data structure.

We add terms representing operations that create a vector given a list of terms, 
get a reference (mutable or immutable) of an element of a vector given a reference to the vector and an index, 
a similar method for pushing an element to a vector, and a method for popping an element
(and returning the actual element as opposed to just a reference of it). Any term can also be given a type annotation.
Note that in this implementation, all variables are mutable.

@@

Lastly, when relevant, our typing rules make use of the following named set,
denoting types that are copyable.
$$\text{Copy}=\{\text{Nat}, \text{Bool}, \text{Unit},{}_{x}\&^\kappa_{\text{shr}} T\}$$

@@

This pass also does all the necessary work for lifetimes; assigning lifetimes
to each of the variables and their respective borrows (see Figure \ref{varlifetimes}), 
checking the subtype relations for references (T-LftSubtype and T-MutShrSubtype, see Figure \ref{borrowrules}), and handling instantiation of functions polymorphic over
lifetimes (Figure \ref{lifetimepoly}).

\begin{figure}[h]
    \begin{lstlisting}[language=caml]
        let a = 0 in        (* a has lifetime '0 *)
        let b = succ 0 in   (* b has lifetime '1 *)
        let c = (\d. d) in  (* c has lifetime '2 *)
        let d = &b in       (* b has type &'1 nat *)
        let d = &mut a in   (* b has type &'0 mut nat *)
        ...
    \end{lstlisting}

    \caption{Example code illustrating lifetimes of variables}
    \label{varlifetimes}
\end{figure}

\begin{figure}[h]
    \begin{lstlisting}[language=caml]
        let y = (\x. x) : &'a nat -> &'a nat in 
        let k = succ (succ 0) in (* k has lifetime '1 *)
        let z = y &k in          (* z has type &'1 nat*)
        ...
    \end{lstlisting}

    \caption{Example code illustrating lifetime polymorphic instantiation}
    \label{lifetimepoly}
\end{figure}
